Ground Truth Sampling Cells Colorado Plateau

### 2024-04-03 scripts/GroundTruthSampling.Rmd
adding fn to select unique ish cells

```{r}
library(sf)
library(terra)
```

@@

>>>>

```{r}
library(sf)
library(terra)
set.seed(23)
```

@@

```{r mask areas with fewer predicted species}
samples <- spatSample(r, 10000, na.rm = T)

lyr1 <- as.numeric(quantile(samples$SumOccupiedPatches, probs = 0.75))
lyr2 <- as.numeric(quantile(samples$Sum1stOrderNeighboringPatches, probs = 0.75))
lyr3 <- as.numeric(quantile(samples$CommonessInversion, probs = 0.75))

r[[1]] <- ifel(r[[1]] < lyr1, NA, r[[1]])
r[[2]] <- ifel(r[[2]] < lyr2, NA, r[[2]])
r[[3]] <- ifel(r[[3]] < lyr3, NA, r[[3]])

# we will need to sample 2 '1's per species, 3 '2-4's per species.

rm(lyr1, lyr2, lyr3)
```

```{r}
datr <- lapply(dat, rasterize, r, field = 'Rank')
datr <- rast(datr)
```

```{r}

samplesFocal <- vector(mode = 'list', length = dim(datr)[3])
samples1stOrder <- vector(mode = 'list', length = dim(datr)[3])

for (i in seq_along(1:dim(datr)[3])){

  samplesFocal[[i]] <- spatSample(ifel(datr[[i]] == 1, 1, NA), 50, na.rm = TRUE, exhaustive = TRUE, as.points = TRUE)
  samplesFocal[[i]] <- extract(r, samplesFocal[[i]], method = 'bilinear', xy = TRUE)
  samples1stOrder[[i]] <- spatSample(ifel(datr[[i]] %in% c(2, 3, 4), 1, NA), 50, na.rm = TRUE, exhaustive = TRUE, as.points = TRUE)
  samples1stOrder[[i]] <- extract(r, samples1stOrder[[i]], method = 'bilinear', xy = TRUE)
  names(samplesFocal) <- names(datr) ; names(samples1stOrder) <- names(datr)

}

# now we determine which are within X distance of each other... 
ob <- data.frame(
  rbind(
  data.table::rbindlist(samplesFocal, idcol = 'taxon') |>
    dplyr::mutate('Target' = 'Populated'),
  data.table::rbindlist(samples1stOrder, idcol = 'taxon') |>
    dplyr::mutate('Target' = '1stOrder')
)) %>% 
  dplyr::arrange(taxon) %>% 
  sf::st_as_sf(coords = c('x', 'y'), crs = 5070)

ob<- split(ob, f = ob$taxon)

identify_unqiue_pts <- function(x){
  # determine which pairs are unique. 
  DAT <- nngeo::st_nn(x, x, k = 5, maxdist = 2500, sparse = TRUE, returnDist = TRUE)
  names(DAT$nn) <- 1:length(DAT$nn)
  dd <- DAT$nn[ lengths( DAT$nn ) >= 2 ]

  # now we need to remove records which are within this distance from each other... 
  pairs <- vector(mode = 'list', length = length(dd))
  for (i in 1:length(dd)){
    pairs[[i]] <- which ( lapply(dd, setequal, dd[[i]]) == TRUE)
  }

  chosen_pts <- c(
    DAT$nn[ lengths( DAT$nn ) == 1 ],
    DAT$nn[as.numeric(names( unlist(lapply(unique(pairs), sample, 1)) ) )]
  ) |> 
    names() |>
    as.numeric() |>
    sort()

  out <- x[chosen_pts,]
  return(out)

}

lapply(ob, identify_unqiue_pts)
```
